A Generalization of Regularized Discriminant Analysis for High-Dimensional Classification

### 2013-03-25 reports/paper-rda.tex
Added simulation results, a discussion of the results, and a discussion to the paper.

Alternatively, several researchers, including \cite*{Zhang:2010va}, \cite{Ye:2009gd}, \cite{Ji:2008wp}, \cite*{Guo:2007te}, \cite{Srivastava:2007ud}, \cite{Ye:2006vx}, \cite*{Thomaz:2006ef}, \cite{Ye:2005uu}, \cite{Mkhadri:1995jp}, and \cite{Peck:1982tt}, have proposed regularization methods to improve the estimation of the population covariance matrices before applying the \emph{LDA} and \emph{QDA} classifiers. Covariance-matrix regularization methods often are based on the well-known ridge-regression approach from \cite{Hoerl:1970cd} and adjust the eigenvalues of the sample covariance matrices to ensure that the estimator is positive definite. Typically, a positive shrinkage value is added to each eigenvalue estimate, replacing the zero eigenvalues with the shrinkage constant and stabilizing the eigenvalues near zero to ensure that the inverse of the the sample covariance matrix is well-posed. This common form of covariance-matrix regularization is equivalent to summing the original covariance-matrix estimator with the identity matrix scaled by the shrinkage constant. For an excellent overview of the regularization problem in discriminant analysis, see \cite{Murphy:2012uq} and \cite{Mkhadri:1997gy}.

Here, we consider a generalization of the popular \emph{RDA} classifier of \cite{Friedman:1989tm}, who has proposed a weighted average of the covariance-matrix estimators from the \emph{LDA} and \emph{QDA} classifiers. Because the resulting biased covariance-matrix estimator can remain singular or nearly singular, \cite{Friedman:1989tm} has additionally incorporated applied covariance-matrix shrinkage, where the shrinkage constant is a function of the average eigenvalue of the original estimator. \cite{Friedman:1989tm} has shown that the \emph{RDA} classifier can attain improved classification accuracy, compared to the \emph{LDA} and \emph{QDA} classifiers when $p > N$. However, in its original form, the \emph{RDA} classifier is impractical when $p \gg N$ because the \emph{RDA} decision rule requires the calculation of the inverses and determinants of the $p \times p$ covariance-matrix estimators for each class. Furthermore, the \emph{RDA} classifier has two tuning parameters that are selected via cross-validation, which adds a substantial computational burden to the model-selection process. Despite the introduction of a computational shortcut by \cite{Friedman:1989tm} for model selection to alleviate some of the computation, the \emph{RDA} classifier is ill-advised on modern data sets because the model selection requires on the order of $Kp^3$ calculations for each tuning-parameter pair for each fold in the cross-validation scheme.

@@

In addition to the improvement in computational performance of classifiers with regularized covariance matrices, we demonstrate that our \emph{GRDA} classifier is competitive with and often superior to several recent classifiers designed for small-sample, high-dimensional data. In particular, we compare our \emph{GRDA} classifier with an alternatively-defined \emph{RDA} classifier from \cite*{Hastie:2008dt}, the regularized \emph{LDA} classifier from \cite{Guo:2007te}, a recently proposed variant of the \emph{DLDA} classifier from \cite*{Tong:2012hw}, and the penalized \emph{LDA} classifier from \cite{Witten:2011kc}. We conduct our comparison of the classification accuracy of the competing classifiers using three small-sample, high-dimensional microarray data sets and examine the classification accuracy of the competing classifiers using \textbf{three} simulation configurations.

We have organized the remainder of the paper as follows. In Section 2 we present our generalization of the \emph{RDA} classifier after reviewing its original counterpart from \cite{Friedman:1989tm}. In Section 3 we show that our proposed classifier includes several well-known regularized classifiers as special cases and briefly discuss some commonalities among them. In Section 4 we describe the \textbf{three} simulation configurations and three small-sample, high-dimensional microarray data sets and then demonstrate that our two proposed classifiers are superior to the competing classifiers in terms of classification accuracy. We then conclude with a brief discussion in Section 5.

@@

In discriminant analysis we wish to assign correctly an unlabeled $p$-dimensional observation vector $\bm x$ to one of $K$ unique, known classes (or populations) by constructing a classifier from $n$ training observations that can accurately predict the class membership of $\bm x$. Let $\bm x_i = (x_{i1}, x_{i2}, \ldots, x_{ip}) \in \mathbb{R}_{p \times 1}$ be the $i$th observation $(i = 1, \ldots, n)$ with true, unique membership $y_i \in \{\omega_1, \ldots, \omega_K\}$, where $\mathbb{R}_{a \times b}$ denotes the matrix space of all $a \times b$ matrices over the real field $\mathbb{R}$. We assume that $(\bm x_i, y_i)$ is a realization from a mixture distribution $p(\bm x) = \sum_{k=1}^K p(\bm x | \omega_k) p(\omega_k)$, where $p(\bm x | \omega_k)$ is the probability density function (PDF) of the $k$th class, $p(\omega_k)$ is prior probability of class membership of the $k$th class.

The \emph{QDA} classifier is the optimal Bayesian decision rule with respect to a $0-1$ loss function when $p(\bm x | \omega_k)$ is the PDF of the multivariate normal distribution with known mean vector  $\bm\mu_k \in \mathbb{R}_{p \times 1}$ and known covariance matrix $\bm\Sigma_k \in \mathbb{R}_{p \times p}^{>}$, $k = 1, 2, \ldots, K$, where $\mathbb{R}_{p \times p}^{>}$ denotes the cone of real $p \times p$ positive-definite matrices. Because the parameters are typically unknown, we estimate the unknown parameters $\bm \mu_k$ and $\bm\Sigma_k$ with their maximum likelihood estimators (MLEs) and substitute the MLEs into the \emph{QDA} classifier. Assuming that the prior probabilities of class membership $p(\omega_k)$ are equal for $k = 1, \ldots, K$, we assign an unlabeled observation $\bm x$ to class $\omega_k$ with the sample \emph{QDA} classifier
\begin{align}
	D_{QDA}(\bm x) = \argmin_{k}  (\bm x - \xbar_k)'\widehat{\bm\Sigma}_k^{-1}(\bm x - \xbar_k)  + \ln |\widehat{\bm\Sigma}_k|, \label{eq:qda}
\end{align}
where $\xbar_k$ and $\widehat{\bm\Sigma}_k$ are the MLEs for $\bm \mu_k$ and $\bm \Sigma_k$, respectively. If we assume further that the population covariance matrices are equal for each class (i.e., $\bm\Sigma_k = \bm\Sigma$ for all $k$), then the pooled sample covariance matrix $\widehat{\bm\Sigma}$ is substituted for $\widehat{\bm \Sigma}_k$ in \eqref{eq:qda}, where
\begin{align}
	\widehat{\bm\Sigma} = \frac{1}{n} \sum_{k=1}^K n_k \widehat{\bm\Sigma}_k. \label{eq:pooled-cov}
\end{align}
is the MLE for $\bm \Sigma$. In this case \eqref{eq:qda} reduces to Fisher's \emph{LDA} classifier, and the log-determinant is omitted because it is constant across all classes.

@@

Several regularization methods, such as the methods considered by \citet*{Guo:2007te}, \cite{Mkhadri:1995jp}, and \citet*{Xu:2009fl}, have been proposed in the literature to adjust the eigenvalues of $\widehat{\bm \Sigma}_k$ so that \eqref{eq:qda} is calculable and has reduced variability in the standard bias-variance tradeoff.  A common form of the covariance-matrix regularization applies a shrinkage factor $\gamma_k > 0$, so that
\begin{align}
	\widehat{\bm \Sigma}_k(\gamma_k) =  \widehat{\bm \Sigma}_k + \gamma_k \bm I_p, \label{eq:ridge-estimator}
\end{align}
similar to a method employed in ridge regression \citep{Hoerl:1970cd}. Equation \eqref{eq:ridge-estimator} effectively \emph{shrinks} the sample covariance matrix $\widehat{\bm\Sigma}_k$ towards $\bm I_p$, thereby increasing the eigenvalues of $\widehat{\bm\Sigma}_k$ by $\gamma_k$. Specifically, the zero eigenvalues are replaced with $\gamma_k$, so that \eqref{eq:ridge-estimator} is positive definite. For additional covariance-matrix regularization methods, see \cite{Ramey:2011ji}, \cite{Xu:2009fl}, and \cite{Ye:2009gd}.

@@

\cite{Friedman:1989tm} has proposed the \emph{RDA} classifier by incorporating a weighted average of the sample covariance matrix $\widehat{\bm \Sigma}_k$ for class $\omega_k$ and the pooled sample covariance matrix $\widehat{\bm\Sigma}$ to estimate the covariance matrix for class $\omega_k$ with
\begin{align}
  \widehat{\bm\Sigma}_k(\lambda) = n_k^{-1}(\lambda) \bm S_k(\lambda),\label{eq:sig-lambda}
\end{align}
where $\lambda \in [0, 1]$, $n_k(\lambda) = (1 - \lambda) n_k + \lambda n$, $\bm S_k = n_k \widehat{\bm\Sigma}_k$, $\bm S = \sum_{k=1}^K \bm S_k$, and $\bm S_k(\lambda) = (1 - \lambda) \bm S_k + \lambda \bm S$. We can interpret \eqref{eq:sig-lambda} as a covariance matrix estimator for class $\omega_k$ that borrows from $\widehat{\bm\Sigma}$ in \eqref{eq:pooled-cov} to better estimate $\bm \Sigma_k$. If $\lambda = 0$, \eqref{eq:sig-lambda} is equal to $\widehat{\bm\Sigma}_k$ used in the \emph{QDA} classifier in \eqref{eq:qda}. Alternatively, if $\lambda = 1$, \eqref{eq:sig-lambda} reduces to \eqref{eq:pooled-cov} used in the \emph{LDA} classifier.

@@

In this section, we describe \textbf{three} high-dimensional microarray data sets and compare our proposed \emph{GRDA} classifier with \textbf{three} recently proposed classifiers for small-sample, high-dimensional data: the penalized \emph{LDA} (\emph{PLDA}) classifier from \cite{Witten:2011kc}, the shrinkage-mean-based DLDA (\emph{SmDLDA}) classifier from \cite{Tong:2012hw}, and the sparse discriminant analysis (\emph{SDA}) classifier from \cite{Clemmensen:2011kr}. We evaluated the classification performance of each classifier by randomly partitioning the data such that $2/3$ of the observations were allocated as training data and the remaining $1/3$ of the observations were allocated as a test data set. For each random partition, we applied the feature-selection method from \cite{Dudoit:2002ev} to choose the $d$ features having the largest ratio of their between-group to within-group sums of squares from the training data. We then filtered the test data set with the corresponding $d$ features. After training each classifier from the filtered training data, we classified the filtered test data sets and computed the proportion of mislabeled test observations to estimate the classification error rate for each classifier. Repeating this process 50 times, we computed the average of the error-rate estimates for each classifier. We remark that our simulation design avoids the selection bias discussed in \cite{Ambroise:2002fa}.

We used version 2.15.1 of the open-source statistical software {\tt R} for the classification study. For each classifier's {\tt R} implementation, we used the default settings with the exception that we explicitly set the prior probabilities as equal. In our {\tt R} implementation of the \emph{GRDA} classifier in the {\tt regdiscrim} package, we set $\alpha_k = 1$, $k = 1, \ldots, K$ for simplicity. We estimated $\lambda$ from a grid of 21 equidistant candidate values between 0 and 1, inclusively. Similarly, we selected $\gamma$ from a grid consisting of the values $10^{-1}, \ldots, 10^4$, and $10^5$. We selected optimal estimates of $\lambda$ and $\gamma$ using 10-fold cross-validation \citep{Hastie:2008dt}.

@@

@@

\cite{Singh:2002fh} have examined 235 radical prostatectomy specimens from surgery patients between 1995 and 1997. The authors used oligonucleotide microarrays containing probes for approximately 12,600 genes and expressed sequence tags. They have reported that 102 of the radical prostatectomy specimens are of high quality: 52 prostate tumour samples and 50 non-tumour prostate samples.

@@

As we can see, our competitors attain decent classification accuracy only when the number of variables selected is small. As the number of variables is increased, the classification accuracy of our competitors degrades.

@@

Our formulation of covariance-matrix regularization employed in the \emph{GRDA} classifier includes a large family of shrinkage methods, such as the methods from \cite{Srivastava:2007ww}, \cite{Rao:1971ul}, and a wide variety of other shrinkage methods studied by \cite{Ramey:2011ji} and \cite*{Xu:2009fl}.

@@

>>>>

Alternatively, several researchers, including \cite*{Zhang:2010va}, \cite{Ye:2009gd}, \cite{Ji:2008wp}, \cite*{Guo:2007te}, \cite{Srivastava:2007ud}, \cite{Ye:2006vx}, and \cite{Ye:2005uu}, have proposed regularization methods to improve the estimation of the population covariance matrices before applying the \emph{LDA} and \emph{QDA} classifiers. Covariance-matrix regularization methods often are based on the well-known ridge-regression approach from \cite{Hoerl:1970cd} and adjust the eigenvalues of the sample covariance matrices to ensure positive definiteness. For excellent overviews of the regularization problem in discriminant analysis, see \cite{Murphy:2012uq} and \cite{Mkhadri:1997gy}.

Here, we consider a generalization of the popular \emph{RDA} classifier of \cite{Friedman:1989tm}, who has proposed a weighted average of the covariance-matrix estimators from the \emph{LDA} and \emph{QDA} classifiers. \cite{Friedman:1989tm} has additionally incorporated a covariance-matrix shrinkage term to ensure that the \emph{RDA} classifier is well-posed. \cite{Friedman:1989tm} has shown that the \emph{RDA} classifier can attain improved classification accuracy, compared to the \emph{LDA} and \emph{QDA} classifiers when $p > N$. However, in its original form, the \emph{RDA} classifier is impractical when $p \gg N$ because the \emph{RDA} decision rule requires the calculation of the inverses and determinants of the $p \times p$ covariance-matrix estimators for each class. Furthermore, the \emph{RDA} classifier has two tuning parameters that are selected via cross-validation, which adds a substantial computational burden to the model-selection process. The \emph{RDA} classifier is ill-advised on modern data sets because the model selection requires $O(Kp^3)$ calculations for each tuning-parameter pair for each fold in the cross-validation scheme.

@@

In addition to the improvement in computational performance of classifiers with regularized covariance matrices, we demonstrate that our \emph{GRDA} classifier is competitive with and often superior to several recent classifiers designed for small-sample, high-dimensional data. In particular, we compare our \emph{GRDA} classifier with an alternatively-defined \emph{RDA} classifier from \cite*{Hastie:2008dt}, the regularized \emph{LDA} classifier from \cite{Guo:2007te}, a recently proposed variant of the \emph{DLDA} classifier from \cite*{Tong:2012hw}, and the penalized \emph{LDA} classifier from \cite{Witten:2011kc}. We conduct our comparison of the classification accuracy of the competing classifiers using four small-sample, high-dimensional microarray data sets.

We have organized the remainder of the paper as follows. In Section 2 we present our generalization of the \emph{RDA} classifier after reviewing its original counterpart from \cite{Friedman:1989tm}. In Section 3 we show that our proposed classifier includes several well-known regularized classifiers as special cases and briefly discuss some commonalities among them. In Section 4 we describe four small-sample, high-dimensional microarray data sets and then demonstrate that our two proposed classifiers are superior to the competing classifiers in terms of classification accuracy. We then conclude with a brief discussion in Section 5.

@@

In discriminant analysis we wish to assign correctly an unlabeled $p$-dimensional observation vector $\bm x$ to one of $K$ unique, known classes (or populations) by constructing a classifier from $n$ training observations that can accurately predict the class membership of $\bm x$. Let $\bm x_i = (x_{i1}, \ldots, x_{ip}) \in \mathbb{R}_{p \times 1}$ be the $i$th observation $(i = 1, \ldots, N)$ with true, unique membership $y_i \in \{\omega_1, \ldots, \omega_K\}$, where $\mathbb{R}_{a \times b}$ denotes the matrix space of all $a \times b$ matrices over the real field $\mathbb{R}$. We assume that $(\bm x_i, y_i)$ is a realization from a mixture distribution $p(\bm x) = \sum_{k=1}^K p(\bm x | \omega_k) p(\omega_k)$, where $p(\bm x | \omega_k)$ is the probability density function (PDF) of the $k$th class, $p(\omega_k)$ is prior probability of class membership of the $k$th class.

The \emph{QDA} classifier is the optimal Bayesian decision rule with respect to a $0-1$ loss function when $p(\bm x | \omega_k)$ is the PDF of the multivariate normal distribution with known mean vector  $\bm\mu_k \in \mathbb{R}_{p \times 1}$ and known covariance matrix $\bm\Sigma_k \in \mathbb{R}_{p \times p}^{>}$, $k = 1, 2, \ldots, K$, where $\mathbb{R}_{p \times p}^{>}$ denotes the cone of real $p \times p$ positive-definite matrices. Because the parameters are typically unknown, we estimate the unknown parameters $\bm \mu_k$ and $\bm\Sigma_k$ with their maximum likelihood estimators (MLEs) and substitute the MLEs into the \emph{QDA} classifier. Assuming that the prior probabilities of class membership $p(\omega_k)$ are equal for $k = 1, \ldots, K$, we assign an unlabeled observation $\bm x$ to class $\omega_k$ with the sample \emph{QDA} classifier
\begin{align}
	D_{QDA}(\bm x) = \argmin_{k}  (\bm x - \xbar_k)'\widehat{\bm\Sigma}_k^{-1}(\bm x - \xbar_k)  + \ln |\widehat{\bm\Sigma}_k|, \label{eq:qda}
\end{align}
where $\xbar_k$ and $\widehat{\bm\Sigma}_k$ are the MLEs for $\bm \mu_k$ and $\bm \Sigma_k$, respectively. If we assume further that the population covariance matrices are equal for each class (i.e., $\bm\Sigma_k = \bm\Sigma$ for all $k$), then the pooled sample covariance matrix $\widehat{\bm\Sigma}$ is substituted for $\widehat{\bm \Sigma}_k$ in \eqref{eq:qda}, where
\begin{align}
	\widehat{\bm\Sigma} = \frac{1}{N} \sum_{k=1}^K n_k \widehat{\bm\Sigma}_k. \label{eq:pooled-cov}
\end{align}
is the MLE for $\bm \Sigma$. In this case \eqref{eq:qda} reduces to Fisher's \emph{LDA} classifier, and the log-determinant is omitted because it is constant across all classes.

@@

Several regularization methods, such as the methods considered by \citet*{Guo:2007te}, \cite{Mkhadri:1995jp}, and \citet*{Xu:2009fl}, have been proposed in the literature to adjust the eigenvalues of $\widehat{\bm \Sigma}_k$ so that \eqref{eq:qda} is calculable and has reduced variability in the standard bias-variance tradeoff.  A common form of the covariance-matrix regularization applies a shrinkage factor $\gamma_k > 0$, so that
\begin{align}
	\widehat{\bm \Sigma}_k(\gamma_k) =  \widehat{\bm \Sigma}_k + \gamma_k \bm I_p, \label{eq:ridge-estimator}
\end{align}
similar to a method employed in ridge regression \citep{Hoerl:1970cd}. Equation \eqref{eq:ridge-estimator} effectively \emph{shrinks} the sample covariance matrix $\widehat{\bm\Sigma}_k$ towards $\bm I_p$, thereby increasing the eigenvalues of $\widehat{\bm\Sigma}_k$ by $\gamma_k$. Specifically, the zero eigenvalues are replaced with $\gamma_k$, so that \eqref{eq:ridge-estimator} is positive definite. For additional covariance-matrix regularization methods, see \cite{Ramey:2013ji}, \cite{Xu:2009fl}, and \cite{Ye:2009gd}.

@@

\cite{Friedman:1989tm} has proposed the \emph{RDA} classifier by incorporating a weighted average of the sample covariance matrix $\widehat{\bm \Sigma}_k$ for class $\omega_k$ and the pooled sample covariance matrix $\widehat{\bm\Sigma}$ to estimate the covariance matrix for class $\omega_k$ with
\begin{align}
  \widehat{\bm\Sigma}_k(\lambda) = n_k^{-1}(\lambda) \bm S_k(\lambda),\label{eq:sig-lambda}
\end{align}
where $\lambda \in [0, 1]$, $n_k(\lambda) = (1 - \lambda) n_k + \lambda N$, $\bm S_k = n_k \widehat{\bm\Sigma}_k$, $\bm S = \sum_{k=1}^K \bm S_k$, and $\bm S_k(\lambda) = (1 - \lambda) \bm S_k + \lambda \bm S$. We can interpret \eqref{eq:sig-lambda} as a covariance matrix estimator for class $\omega_k$ that borrows from $\widehat{\bm\Sigma}$ in \eqref{eq:pooled-cov} to better estimate $\bm \Sigma_k$. If $\lambda = 0$, \eqref{eq:sig-lambda} is equal to $\widehat{\bm\Sigma}_k$ used in the \emph{QDA} classifier in \eqref{eq:qda}. Alternatively, if $\lambda = 1$, \eqref{eq:sig-lambda} reduces to \eqref{eq:pooled-cov} used in the \emph{LDA} classifier.

@@

In this section, we describe four high-dimensional microarray data sets and compare our proposed \emph{GRDA} classifier with five recently proposed classifiers designed for small-sample, high-dimensional data: the \emph{DLDA} classifier from \cite{Dudoit:2002ev}, the shrunken centroids \emph{RDA} (\emph{SCRDA}) classifier from \cite{Guo:2007te}, the shrinkage-based \emph{DLDA} (\emph{SDLDA}) classifier from \cite{Pang:2009ik}, and the shrinkage-mean-based DLDA (\emph{SmDLDA}) classifier from \cite{Tong:2012hw}, and the penalized \emph{LDA} (\emph{PLDA}) classifier from \cite{Witten:2011kc}. We evaluated the classification performance of each classifier by randomly partitioning the data such that $2/3$ of the observations were allocated as training data and the remaining $1/3$ of the observations were allocated as a test data set. For each random partition, we applied the feature-selection method from \cite{Dudoit:2002ev} to choose the $d$ features having the largest ratio of their between-group to within-group sums of squares from the training data. We then filtered the test data set with the corresponding $d$ features. After training each classifier from the filtered training data, we classified the filtered test data sets and computed the proportion of mislabeled test observations to estimate the classification error rate for each classifier. Repeating this process 500 times, we computed the average of the error-rate estimates for each classifier. We remark that our simulation design avoids the selection bias discussed in \cite{Ambroise:2002fa}.  In Table \ref{tab:data-summary}, we summarize the attributes of each microarray data set.

We used version 2.15.1 of the open-source statistical software {\tt R} for the classification study. For each classifier's {\tt R} implementation, we used the default settings with the exception that we explicitly set the prior probabilities as equal. In our {\tt R} implementation of the \emph{GRDA} classifier in the {\tt regdiscrim} package, we set $\alpha_k = 1$, $k = 1, \ldots, K$ for simplicity. We estimated $\lambda$ from a grid of 21 equidistant candidate values between 0 and 1, inclusively. Similarly, we estimated $\gamma$ from a grid consisting of the values $10^{-1}, \ldots, 10^4$, and $10^5$. We selected optimal estimates of $\lambda$ and $\gamma$ using 10-fold cross-validation \citep{Hastie:2008dt}.

 % latex table generated in R 2.15.1 by xtable 1.7-0 package
 % Sun Mar 24 19:54:37 2013
 \begin{table}[ht]
 \begin{center}
 \begin{tabular}{lrrrl}
   \hline
 Author & $N$ & $p$ & $K$ & Disease \\
   \hline
 Burczynski et al. (2006) & 127 & 22283 & 3 & Crohn's Disease \\
   Nakayama et al. (2001) & 105 & 22283 & 5 & Sarcoma \\
   Shipp et al. (2002) & 58 & 6817 & 2 & Lymphoma \\
   Singh et al. (2002) & 102 & 12600 & 2 & Prostate Cancer \\
    \hline
 \end{tabular}
 \caption{Summary of high-dimensional microarray data sets.}
 \label{tab:data-summary}
 \end{center}
 \end{table}

@@

\subsection{\cite{Shipp:2002ka} Data Set}

Approximately 30-40\% of adult non-Hodgkin lymphomas are diffuse large B-cell lymphomas (DLBCLs) \cite{Shipp:2002ka}.  However, only a small proportion of DLBCL patients are cured with modern chemotherapeutic regimens.  Several models have been proposed, such as the International Prognostic Index (IPI), to determine a patient's curability.  These models rely on clinical covariates, such as age, to determine if the patient can be cured and are often ineffective. \cite{Shipp:2002ka} have argued that more effective means are desired to determine a patient's curability. The authors measured 6,817 gene expression levels from 58 DLBCL patient samples with customized cDNA (‘lymphochip’) microarrays to investigate the curability of patients treated with cyclophosphamide, adriamycin, vincristine, and prednisone (CHOP)-based chemotherapy.  Among the 58 DLBCL patient samples, 32 are from cured patients while 26 are from patients with fatal or refractory disease.

@@

\cite{Singh:2002fh} have examined 235 radical prostatectomy specimens from surgery patients between 1995 and 1997. The authors used oligonucleotide microarrays containing probes for approximately 12,600 genes and expressed sequence tags. They have reported that 102 of the radical prostatectomy specimens are of high quality: 52 prostate tumor samples and 50 non-tumor prostate samples.

@@

For the microarray data sets, we computed error-rate estimates for the competing classifiers after reducing the original feature space dimension with variable selection to $d=100,250,500$, and $1000$.  Our choices for the values of $d$ are similar to those of \cite{Xu:2009fl}.  We report the error-rate estimates along with approximate standard errors in parentheses in Tables \ref{tab:results-d100} -- \ref{tab:results-d1000} along with the chosen reduced dimension $d$.

The \emph{GRDA} classifier was superior in classification performance for the majority of the simulations. While the error-rate estimates for the majority of the classifiers increased as $d$ was increased, the classification accuracy for the \emph{GRDA} classifier improved as $d$ increased for the Burczynski and Nakayama data sets. For the Nakayama data set, the \emph{GRDA} classifier was outperformed by the Dudoit, Pang, Tong, and Witten classifiers for small values of $d$ but attained classification accuracy comparable to its competitors for larger values of $d$. Moreover, the only situation where the other classifiers considered were competitive with the \emph{GRDA} classifier was in the case of the Nakayama data set.

The error-rate estimates for the Dudoit, Pang, Tong, and Witten classifiers were comparable regardless of the choice of $d$. In particular, the error-rate estimates for Dudoit's and Tong's methods were nearly identical, which suggests that the improved mean estimator employed by Tong provides little improvement to the classification performance in the \emph{DLDA} classifier. The popular classifier from Guo is competitive only for large values of $d$ for the Shipp and Singh data sets. For the Burczynski and Nakayama data sets, the implementation of Guo's method in the {\tt rda} package often classified all of the test observations into a single class, thus yielding larger error rates than its competitors.

\[ \left[\text{Table 2 goes about here }\right]\]

\[ \left[\text{Table 3 goes about here }\right]\]

\[ \left[\text{Table 4 goes about here }\right]\]

\[ \left[\text{Table 5 goes about here }\right]\]

@@

The analysis of high-dimensional microarray data sets is important to modern biomedical studies and presents challenges with supervised classification because the number of gene expressions greatly outnumbers the number of collected responses. Consequently, regularization techniques are often employed to stabilize popular classifiers for improved classification performance. We have provided a generalization of the well-known \emph{RDA} classifier from \cite{Friedman:1989tm}, aimed for small-sample, high-dimensional classification. The covariance-matrix regularization employed in the \emph{GRDA} classifier includes a large family of shrinkage methods. Our formulation of the \emph{GRDA} classifier in terms of the nonzero eigenvalues of the pooled sample covariance matrix indicates that eigenvalue-adjustment methods applied to the \emph{LDA} classifier explains the similiarities in classification accuracy of several regularization methods applied to the \emph{LDA} classifier reported by \cite{Ramey:2013ji} in a high-dimensional classification study.

We have provided an implementation of the \emph{GRDA} classifier in the {\tt regdiscrim} R package, available on CRAN. Our implementation facilitates a rapid classification of high-dimensional data by employing the novel results provided in this paper. Moreover, our 10-fold cross-validation procedure typically requires a couple of seconds for data sets having large $p$. When preliminary variable selection is applied to the data as we performed in our classification study, the time to perform model selection is nearly instantaneous, confirming the computational complexity reported regarding the operations employed in the \emph{GRDA} classifier.

@@

\clearpage

% latex table generated in R 2.15.1 by xtable 1.7-0 package
% Mon Mar 25 07:04:20 2013
\begin{table}[ht]
\begin{center}
\begin{tabular}{lllllll}
  \hline
Dataset & Dudoit & GRDA & Guo & Pang & Tong & Witten \\ 
  \hline
Burczynski & 0.232 (0.019) & 0.179 (0.017) & 0.513 (0.022) & 0.232 (0.019) & 0.231 (0.019) & 0.232 (0.019) \\ 
  Nakayama & 0.262 (0.020) & 0.308 (0.021) & 0.586 (0.022) & 0.262 (0.020) & 0.262 (0.020) & 0.278 (0.020) \\ 
  Shipp & 0.136 (0.015) & 0.106 (0.014) & 0.310 (0.021) & 0.136 (0.015) & 0.136 (0.015) & 0.136 (0.015) \\ 
  Singh & 0.111 (0.014) & 0.097 (0.013) & 0.338 (0.021) & 0.111 (0.014) & 0.111 (0.014) & 0.111 (0.014) \\ 
   \hline
\end{tabular}
\caption{The average of the test error rates for the microarray data sets with $d = 100$. Approximate standard errors are given in parentheses.}
\label{tab:results-d100}
\end{center}
\end{table}

% latex table generated in R 2.15.1 by xtable 1.7-0 package
% Mon Mar 25 07:04:37 2013
\begin{table}[ht]
\begin{center}
\begin{tabular}{lllllll}
  \hline
Dataset & Dudoit & GRDA & Guo & Pang & Tong & Witten \\ 
  \hline
Burczynski & 0.244 (0.019) & 0.163 (0.017) & 0.459 (0.022) & 0.244 (0.019) & 0.244 (0.019) & 0.243 (0.019) \\ 
  Nakayama & 0.244 (0.019) & 0.267 (0.020) & 0.498 (0.022) & 0.244 (0.019) & 0.244 (0.019) & 0.259 (0.020) \\ 
  Shipp & 0.153 (0.016) & 0.104 (0.014) & 0.188 (0.017) & 0.153 (0.016) & 0.153 (0.016) & 0.154 (0.016) \\ 
  Singh & 0.135 (0.015) & 0.117 (0.014) & 0.188 (0.017) & 0.135 (0.015) & 0.135 (0.015) & 0.135 (0.015) \\ 
   \hline
\end{tabular}
\caption{The average of the test error rates for the microarray data sets with $d = 250$. Approximate standard errors are given in parentheses.}
\label{tab:results-d250}
\end{center}
\end{table}

% latex table generated in R 2.15.1 by xtable 1.7-0 package
% Mon Mar 25 07:04:48 2013
\begin{table}[ht]
\begin{center}
\begin{tabular}{lllllll}
  \hline
Dataset & Dudoit & GRDA & Guo & Pang & Tong & Witten \\ 
  \hline
Burczynski & 0.243 (0.019) & 0.159 (0.016) & 0.429 (0.022) & 0.243 (0.019) & 0.243 (0.019) & 0.243 (0.019) \\ 
  Nakayama & 0.247 (0.019) & 0.248 (0.019) & 0.465 (0.022) & 0.247 (0.019) & 0.248 (0.019) & 0.261 (0.020) \\ 
  Shipp & 0.169 (0.017) & 0.097 (0.013) & 0.119 (0.015) & 0.169 (0.017) & 0.169 (0.017) & 0.170 (0.017) \\ 
  Singh & 0.174 (0.017) & 0.125 (0.015) & 0.155 (0.016) & 0.174 (0.017) & 0.174 (0.017) & 0.174 (0.017) \\ 
   \hline
\end{tabular}
\caption{The average of the test error rates for the microarray data sets with $d = 500$. Approximate standard errors are given in parentheses.}
\label{tab:results-d500}
\end{center}
\end{table}

% latex table generated in R 2.15.1 by xtable 1.7-0 package
% Mon Mar 25 07:05:01 2013
\begin{table}[ht]
\begin{center}
\begin{tabular}{lllllll}
  \hline
Dataset & Dudoit & GRDA & Guo & Pang & Tong & Witten \\ 
  \hline
Burczynski & 0.249 (0.019) & 0.168 (0.017) & 0.381 (0.022) & 0.249 (0.019) & 0.249 (0.019) & 0.249 (0.019) \\ 
  Nakayama & 0.237 (0.019) & 0.238 (0.019) & 0.460 (0.022) & 0.237 (0.019) & 0.237 (0.019) & 0.246 (0.019) \\ 
  Shipp & 0.196 (0.018) & 0.106 (0.014) & 0.095 (0.013) & 0.196 (0.018) & 0.195 (0.018) & 0.198 (0.018) \\ 
  Singh & 0.224 (0.019) & 0.140 (0.016) & 0.142 (0.016) & 0.224 (0.019) & 0.224 (0.019) & 0.224 (0.019) \\ 
   \hline
\end{tabular}
\caption{The average of the test error rates for the microarray data sets with $d = 1000$. Approximate standard errors are given in parentheses.}
\label{tab:results-d1000}
\end{center}
\end{table}
